---
title: "544 project"
date: "2023-03-11"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r include=FALSE}
options(tinytex.verbose = TRUE)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(sf)
library(SUMMER)
library(tidyverse)
library(spdep)
library(INLA)
library(haven)
library(survey)
library(janitor)
library(sjPlot)

setwd("~/Desktop/554_project/data")
# Population base: De facto children age 6-59 months (PR file)
data <- read_dta("KEPR81FL.DTA")
```


##-----------------------------------Data Preparation----------------------------##

```{r}
data$hc1
#create malaria variables
data<-data%>%mutate(malaria=case_when(hml35 == 1|hml32 == 1 ~ 1,
                                      hml35 == 0 & hml32 == 0 ~ 0),
                    malaria_test =ifelse(malaria %in% c(1,0),1,0))

breaks <- c(0, 12, 24, 36, 48,60)
labels <- c("0-12", "13-24", "25-36", "37-48", "49-60")
data$age_category <- factor(cut(data$hc1, breaks = breaks, labels = labels))


data <-data%>%dplyr::select(hv001,hv002,hv004,hv005,hv021,hv022,hv025,age_category,hc27,hml35,hml32,malaria,malaria_test)

# data prep
data <- data %>%
  mutate(clustid = hv001, 
         id = hv002,
         weight = hv005,
         strata = hv022,
         Sex = factor(ifelse(hc27 == 1, "Male", "Female")),
         urban = ifelse(hv025 == 1, 1, 0)) %>%
  dplyr::select(clustid, id, weight, strata, Sex, age_category,urban,malaria_test, malaria)

head(data)

# load gadm shapefile and add numeric region ID based on admin 2
poly_adm2<- st_read(dsn = ".", layer = "gadm41_KEN_1")
#poly_adm2 <- readRDS("~/Desktop/554_project/data/Subcounty Kenya/gadm36_KEN_2_sp.rds")
poly_adm2$region <- 1:nrow(poly_adm2)

# load GPS data from DHS
points <- st_read(dsn=".", layer="KEGE81FL")

# detect points in the DHS GPS file with mis-specified coordinates and remove them if any
wrong.points <- which(points$LATNUM == 0.0 & points$LONGNUM == 0.0)
if (length(wrong.points) > 0) message("Removing ", length(wrong.points), " clusters with bad GPS data.")
data <- data[!(data$clustid %in% points$DHSCLUST[wrong.points]),]

# merge cluster GPS coordinates onto cluster data then GADM location names onto data using coordinates
data <- merge(points, data, by.y = "clustid", by.x = "DHSCLUST")
names(data)[names(data) == "DHSCLUST"] <- "clustid"
data <- st_join(data, poly_adm2, join = st_nearest_feature, left = T)

data<-data%>%dplyr::select(clustid, id,weight, strata, NAME_1,urban, Sex, age_category,malaria_test, malaria, region)


```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# create adjacency matrix
library(spdep)
library(prioritizr)


# create neighborhood matrix for later
admin2_mat <- poly2nb(poly_adm2)
nb2INLA(file = "admin2_lbr.graph", admin2_mat)
admin2_mat <- nb2mat(admin2_mat, zero.policy = TRUE)
colnames(admin2_mat) <- rownames(admin2_mat) <- poly_adm2$NAME_1
admin2_mat[1:10, 1:10]


nc_df2 <- data %>% st_drop_geometry()


library(raster)
Kenya2<-getData("GADM", country="KE", level=1)

Kenya2@data

Kenya2 <- sp::merge(x = Kenya2, 
                     y = poly_adm2[ , c("NAME_1", "region")], 
                     by.x = c("NAME_1"), 
                     by.y = c("NAME_1"))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Map of Kenya with county name- figure 1
data(Kenya2)
library(randomcoloR)
n <- 47
palette <- distinctColorPalette(n)
                 
ggplot(data = poly_adm2) + geom_sf(aes(fill = NAME_1)) +labs(title = "Map for Kenya") + scale_fill_discrete(name = "County Name") + theme( panel.grid = element_blank(), panel.border = element_blank(), axis.ticks = element_blank(),axis.text = element_blank())
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Map of Kenya with MIS Cluster locations and county boundaries- figure 2
plot(st_geometry(poly_adm2))
# test:
p <- st_as_sf(data)
plot(p, add = TRUE, col = "blue", pch = 15, cex = 0.6)
```


##---------------------------CREATE DATAFRAME FOR BUILDING MODELS-----------------------------##
```{r echo=FALSE, message=FALSE, warning=FALSE}
counties <- c("Mombasa", "Kwale", "Kilifi", "Tana River", "Lamu", "Taita Taveta", "Garissa", "Wajir", "Mandera", "Marsabit", "Isiolo", "Meru", "Tharaka Nithi", "Embu", "Kitui", "Machakos", "Makueni", "Nyandarua", "Nyeri", "Kirinyaga", "Murang'a", "Kiambu", "Turkana", "West Pokot", "Samburu", "Trans Nzoia", "Uasin Gishu", "Elgeyo-Marakwet", "Nandi", "Baringo", "Laikipia", "Nakuru", "Narok", "Kajiado", "Kericho", "Bomet", "Kakamega", "Vihiga", "Bungoma", "Busia", "Siaya", "Kisumu", "Homa Bay", "Migori", "Kisii", "Nyamira","Nairobi")

# create a vector of provinces in Kenya
provinces <- c("Coast", "Coast", "Coast", "Coast", "Coast", "Coast", "North Eastern", "North Eastern", "North Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Eastern", "Central", "Central", "Central", "Central", "Central", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Rift Valley", "Western","Western", "Western", "Western", "Nyanza", "Nyanza", "Nyanza", "Nyanza", "Nyanza", "Nyanza","Nairobi")

# create a dataframe by combining the two vectors
kenya_county_province_df <- data.frame(county = counties, province = provinces)



province<-nc_df2 %>% 
  group_by(NAME_1) %>% 
   summarise(n_cases  = n(),test =sum(malaria_test==1), malaria=sum(malaria_test==1&malaria==1), `Sampled Urban`=sum(urban==1), `Sampled rural`=sum(urban==0), cluster=length(unique(clustid)))%>% as.data.frame()


#Sample allocation of clusters in each region and number of children age 0-60 months interviewed for MIS- table3 supplementary
descriptive <- merge(x = province, y = kenya_county_province_df[ , c("province", "county")], 
                     by.x = c("NAME_1"), 
                     by.y = c("county"))



descriptive<- cbind(descriptive[, ncol(descriptive)], descriptive[, -ncol(descriptive)])
names(descriptive)[1] <- "province"

descriptive%>%dplyr::arrange(province,NAME_1) %>%adorn_totals(name = 'Total')%>%knitr::kable("simple")
#tab_df(province)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# remove records with missing
nc_df2 <- subset(nc_df2, !is.na(data$malaria))
```


## Descriptive statistics with the pooled sample of MIS data in Kenya- Table 1
```{r}
library(table1)

pvalue <- function(x, ...) {
  y <- unlist(x)
  g <- factor(rep(1:length(x), times=sapply(x, length))) 
  if (is.numeric(y)) {
        # For numeric variables, Welch's Two Sample t-test
  p <- t.test(y ~ g)$p.value 
  } else {
        # For categorical variables, Pearson's Chi-squared Test
        p <- chisq.test(table(y, g))$p.value
    }
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
destable<-nc_df2
destable <- merge(x = destable, 
                     y = kenya_county_province_df[ , c("province", "county")], 
                     by.x = c("NAME_1"), 
                     by.y = c("county"))
destable$malaria <- factor(destable$malaria, levels=c(0,1),
         labels=c("Malaria Negative","Malaria Positive"))
destable$urban <- factor(destable$urban, levels=c(0,1),labels=c("Rural", "Urban"))
table1_CMB<-table1(~ age_category+Sex+urban+province| malaria, data = destable, extra.col=list(`P-value`=pvalue))
table1_CMB
knitr::kable(table1_CMB)

```



##-----------------------------BUILDUNG MODELS-------------------------#
## Naive estimates
```{r echo=FALSE, message=FALSE, warning=FALSE}
# compute naive estimates and standard errors- table 4 supplementary
naive_data <- nc_df2 %>%
  group_by(NAME_1) %>%
  summarise(est = mean(malaria),
            n = length(malaria)) %>%
  mutate(se = sqrt( est * (1 - est)/n)) %>%
  as.data.frame()
# mapPlot() doesn't handle tibbles well, so if you're using tidyverse functionality,
# convert your tibbles back to dataframes before plotting
tab_df(naive_data)
naive_data%>%knitr::kable("simple")

# map of estimates
mapPlot(data = naive_data, geo = Kenya2, variables = "est",
    by.data = "NAME_1", by.geo = "NAME_1", legend.label = "Naive Est.",
    is.long = FALSE) +
   ggtitle("Naive Estimates") +
   theme(strip.background = element_blank(),
         plot.title = element_text(hjust = 0.5),
         panel.border = element_blank(),
         strip.text.x = element_blank())


# map of standard errors
mapPlot(data = naive_data, geo = Kenya2, variables = "se",
    by.data = "NAME_1", by.geo = "NAME_1", legend.label = "SE",
    is.long = FALSE) +
   ggtitle("Standard Errors (naive estimates)") +
   theme(strip.background = element_blank(),
         plot.title = element_text(hjust = 0.5),
         panel.border = element_blank(),
         strip.text.x = element_blank())

```











## direct estimate(HT), WEIGHTED ESTIMATE
```{r echo=FALSE, message=FALSE, warning=FALSE}
# compute direct estimates and standard errors-table 5 supplementary
design <- svydesign(ids = ~1, weights = ~weight, strata = ~strata,
                    data = nc_df2)
direct <- svyby(~malaria, ~NAME_1, design, svymean)
head(direct, n = 7)
#tab_df(direct)
direct%>%knitr::kable("simple")


# map of weighted estimates
mapPlot(data = direct, geo = Kenya2, variables = "malaria",
    by.data = "NAME_1", by.geo = "NAME_1", legend.label = "Weighted Est.",
    is.long = FALSE) +
   ggtitle("Weighted Estimates") +
   theme(strip.background = element_blank(),
         plot.title = element_text(hjust = 0.5),
         panel.border = element_blank(),
         strip.text.x = element_blank())


# map of standard errors
mapPlot(data = direct, geo = Kenya2, variables = "se",
    by.data = "NAME_1", by.geo = "NAME_1", legend.label = "SE",
    is.long = FALSE) +
   ggtitle("Standard Errors (weighted estimates)") +
   theme(strip.background = element_blank(),
         plot.title = element_text(hjust = 0.5),
         panel.border = element_blank(),
         strip.text.x = element_blank())

```







## FH-smoothed weighed
```{r echo=FALSE, message=FALSE, warning=FALSE}
par(pty="s")
muD <- mean(direct$malaria)
sdD <- sd(direct$malaria)
qqnorm((direct$malaria-muD)/sdD,main="")
abline(0,1,col="red")

options(survey.adjust.domain.lonely=TRUE)
options(survey.lonely.psu="adjust")

## FH-nonspatial (IID; FHmodel)
FHmodeliid <- smoothSurvey(data = nc_df2, geo = Kenya2, Amat = NULL,
responseType = "binary", responseVar = "malaria", strataVar = "strata",
weightVar = "weight", regionVar = "NAME_1", clusterVar = "~clustid+id",
CI = 0.95)

## FH-spatial (BYM2; svysmoothed)-indirect
FHmodel <- smoothSurvey(data = nc_df2, geo = Kenya2, Amat = admin2_mat,
responseType = "binary", responseVar = "malaria", strataVar = "strata",
weightVar = "weight", regionVar = "NAME_1", clusterVar = "~clustid+id",
CI = 0.95)


# extract posterior medians and map them
toplotFH <- FHmodel$smooth
mapPlot(data = toplotFH, geo = Kenya2, variables = c("median"),
labels = c("Posterior Median"), by.data = "region", by.geo = "NAME_1")+
   ggtitle("Weighted Smoothed Estimates")+
   theme(strip.background = element_blank(),
         plot.title = element_text(hjust = 0.5),
         panel.border = element_blank(),
         strip.text.x = element_blank())

toplotFH$sd <- sqrt(toplotFH$var)
mapPlot(data = toplotFH, geo = Kenya2, variables = c("sd"),labels = c("Posterior sd"),
 by.data = "region", by.geo = "NAME_1")+
  ggtitle("Standard deviation (weighted smoothed estimates)")+
   theme(strip.background = element_blank(),
         plot.title = element_text(hjust = 0.5),
         panel.border = element_blank(),
         strip.text.x = element_blank())


# compute FH estimates and standard errors-table 6 supplementary
FHmodel$smooth[,c(1,2,3,4)]%>% as.data.frame()%>%mutate_if(is.numeric, round, 2)%>% knitr::kable("simple")

```
In supplementary document, provide a quantile-quantile plot comparing the direct estimates to a Gaussian distribution. Here the observed quantiles align well with those from a Gaussian distribution, which lends no support to the basic IID model.






## Table 2 for Malaria Prevalence in each estimation
```{r echo=FALSE, message=FALSE, warning=FALSE}

collect<-data.frame(region = toplotFH$region,
           naive = (naive_data$est * 100) %>% round(1),
           weighted = (direct$malaria * 100) %>% round(1),
           smoothed_weighted = (toplotFH$median * 100) %>% round(1))



collect <- merge(x = collect, 
                     y = kenya_county_province_df[ , c("province", "county")], 
                     by.x = c("region"), 
                     by.y = c("county"))



collect <- cbind(collect[, ncol(collect)], collect[, -ncol(collect)])
names(collect)[1] <- "province"

collect%>%dplyr::arrange(province, smoothed_weighted) %>%knitr::kable("simple")



```



```{r}
## -------------comparing models by scatter plots-----------##
# plot point estimates against each other
data.frame(direct = direct$malaria, naive = naive_data$est) %>%
  ggplot(aes(naive, direct)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  coord_fixed() +
  xlab("Naive estimate") +
  ylab("Weighted estimate")

# plot standard errors against each other
data.frame(direct = direct$se, naive = naive_data$se) %>%
  ggplot(aes(naive, direct)) +
  geom_point() +
  xlab("Naive estimate SE") +
  ylab("Weighted estimate SE") +
  xlim (0, 0.07) +
  ylim (0, 0.07) +
  geom_abline(slope = 1, intercept = 0, col = "red")


# plot point estimates against each other
data.frame(smoothed = toplotFH$median, weighted = direct$malaria) %>%
  ggplot(aes(weighted, smoothed)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  xlab("Weighted estimate") +
  ylab("Smoothed weighted estimate")


# plot standard error and posterior sd against each other
data.frame(smoothed = toplotFH$sd, weighted = direct$se) %>%
  ggplot(aes(weighted, smoothed)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  xlab("Weighted estimate") +
  ylab("Smoothed weighted estimate")

```






\newpage
# Appendix: All code for this report

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
